\section{Governing Equations of Fluid Dynamics}
We first start of with a fluid with a density
\begin{align}
    \rho(\mathbf{x}, t),
\end{align}
in three dimensional Cartesian coordinates $\mathbf{x} = (x, y, z)$ at time
$t$. For water-wave applications, we should note that we take
$\rho=\text{constant}$, but we will go into this fact later. The fluid moves
in time and space with a velocity field
\begin{align}
    \mathbf{u}(\mathbf{x}, t) = (u, v, w).
\end{align}
Additionally it is also described by its pressure
\begin{align}
    P(\mathbf{x}, t),
\end{align}
generally depending on time and position. When thinking of e.g. water the
pressure increases the deeper we go, that is with decreasing or increasing $z$
direction (depending how we set up our system $z$ pointing up or down
respectively).

The general assumption in fluid dynamics is the \textbf{Continuum
Hypothesis}, which assumes continuity of $\textbf{u}, \rho$ and $P$ in
$\mathbf{x}$ and $t$. In other words, we premise that the velocity field,
density and pressure are ''nice enough`` functions of position and time, such
that we can do all the differential operations we desire in the framework of
differential analysis.
\subsection{Mass Conservation}
Our aim is to derive a model of the fluid and its dynamics, with respect to
time and position, in the most general way. This is usually done thinking
of the density of a given fluid, which is a unit mass per unit volume,
intrinsically  an integral representation to derive these equations suggests
by itself.

Let us now thing of an arbitrary fluid. Within this fluid we define a fixed
volume $V$ relative to a chosen inertial frame and bound it by a surface $S$
within the fluid, such that the fluid motion $\mathbf{u}(\mathbf{x}, t)$ may
cross the surface $S$. The fluid density is given by $\rho(\mathbf{x}, t)$,
thereby the mass of the fluid in the defined Volume $V$ is an integral
expression
\begin{align}
    m = \int_V \rho(\mathbf{x}, t) dV.
\end{align}
The figure bellow \ref{fig:volume}, expresses the above described picture.
\begin{figure}[H]
    \centering
  \begin{tikzpicture}[>=latex,scale=1, xscale=1, opacity=.8]
% second sphere
    \begin{scope}[rotate=10, xscale=3, yscale=2, shift={(2.3,-0.2)}]
      \coordinate (O) at (0,0);
      \shade[ball color=gray!10!] (0,0) coordinate(Hp) circle (1) ;

      \draw[thick] (O) circle (1);
      \draw[rotate=5] (O) ellipse (1cm and 0.66cm);
      \draw[rotate=90] (O) ellipse (1cm and 0.33cm);
\node[circle, fill=black, inner sep=1pt] at (0.15, 0.25) {} ; \draw[-latex, thick] (0.15, 0.25) -- (1, 1) ;
      \node[right] at (1, 1) {$\mathbf{u}(\mathbf{x}, t)$};

      \node[] at (O) {$V$};
      \node[] at (0.55, -0.25) {$\rho(\mathbf{x}, t)$};

      \draw[-] (0.76, -0.66) -- (1.2, -0.7);
      \node[right] at (1.2, -0.7) {$S$};

      \draw[-latex, thick] (-0.25, -0.65) -- (-1, -1);
      \node[left] at (-1, -1) {$\mathbf{n}$};

    \end{scope}

% axis
  \end{tikzpicture}
  \caption{Volume bounded by a surface in a fluid with density and momentum,
  with a surface normal vector $\mathbf{n}$ \label{fig:volume}}
\end{figure}

Since we want to figure out the fluid's dynamics, we can consider the rate
of change in the completely arbitrary $V$
\begin{align}
    \frac{d}{dt}\left( \int_V \rho(\mathbf{x}, t)\ dV \right) = \int_V
    \frac{\partial \rho(\mathbf{x}, t)}{\partial t} \ dV
\end{align}
On the other hand we have that density is mass over volume or
\begin{align}
    dm = \rho \cdot dV.
\end{align}
The infinitesimal volume has the base area $dS$ with hight $h$, which is the
distance in the direction perpendicular to to the base area, leaving us with
$dV = h dS$. By definition $\mathbf{n}$ is perpendicular to $dS$, we have
that $h = l \mathbf{n}$. Where $l$ is the unit of length, or velocity times
time $l = \mathbf{u} dt$, since mass is flowing out of the surface we
change the sign of the flow leading us to
\begin{align}
   dm  = -\rho dV = -\rho \mathbf{u} \cdot \mathbf{n} dt dS,
\end{align}
All together we have
\begin{align}
    \frac{dm}{dt} = -\int_{\partial V} \rho(\mathbf{x},t) dS
    \mathbf{u}\cdot\mathbf{n}.
\end{align}
Putting both equations on one side leaves us with a variation of the mass
conservation equation
\begin{align}\label{eq:mass balance}
     \int_V \frac{\partial \rho(\mathbf{x}, t)}{\partial t}\ dV
    +\int_{\partial V} \rho(\mathbf{x}, t) \mathbf{u}\cdot\mathbf{n}\ dS
    = 0.
\end{align}
The above equation in \ref{eq:mass balance} is an underlying equation, describing that the rate of
change of mass in V is brought about, only by the rate of mass flowing into
V across S, and thus the mass does not change.

For the second integral in \ref{eq:mass balance} we utilize the Gaussian
integration law to acquire an integral over the volume
\begin{align}
    \int_{\partial V} \rho(\mathbf{x}, t) \mathbf{u} \cdot \mathbf{n} \ dS =
    \int_V \nabla (\rho \mathbf{u})\ dV.
\end{align}
Thereby we can put everything inside the volume integral
\begin{align}
    \frac{d m}{dt} = \int_V \left(\partial_t \rho + \nabla(\rho \mathbf{u}) \right) \ dV = 0.
\end{align}
Everything under the integral sign needs to be zero, thus we obtain
the \textbf{Equation of Mass Conservation} or in the general sense also
called the \textbf{Continuity Equation}
\begin{align}\label{eq:continuity}
    \partial_t \rho + \nabla(\rho \mathbf{u}) = 0
\end{align}

In light of the results of the equation of mass conservation
in \ref{eq:continuity}, an product rule gives
\begin{align}
    \partial_t \rho + (\nabla \rho)\mathbf{u} + \rho(\nabla \mathbf{u}),
\end{align}
for notational purposes, we define the \textbf{material/convective derivative}
as follows
\begin{align}
    \frac{D}{Dt} = \partial_t + \mathbf{u}\nabla.
\end{align}
With the material derivative the equation of mass conservation reads
\begin{align}
    \frac{D\rho}{Dt} + \rho \nabla\mathbf{u} = 0
\end{align}
We may undertake the first case separation, initiating $\rho = \text{cosnt.}$
called \textbf{incompressible flow} causes the material derivative of $\rho$ to
be zero, and thereby
\begin{align}
    \frac{D\rho}{Dt} = 0 \quad \Rightarrow \quad \nabla \mathbf{u} = 0,
\end{align}
following that the divergence of the velocity field is zero, in this case
$\mathbf{u}$ is called \textbf{solenoidal}.
\subsection{Euler's Equation of Motion}
Additional consideration we undertake is the assumption of an
\textbf{inviscid} fluid, that is we set viscosity to zero. Otherwise we would
get a viscous contribution under the integral which results in the
Navier-Stokes equation. In this regard we apply Newton's second law to our
fluid in terms of infinitesimal pieces $\delta V$ of the fluid. The
acceleration divides into two terms, a \textbf{body force} given by gravity
of earth in the $z$ coordinate $\mathbf{F} = (0, 0, -g)$ and a
\textbf{local/short-rage force} described by the stress tensor in the fluid.
In the inviscid case we the local force retains the pressure $P$, producing a
normal force, with respect to the surface, acting onto any infinitesimal
element in the fluid. The integral formulation of the force would be
\begin{align}
    \int_V \rho \mathbf{F}\ dV - \int_S P\mathbf{n}\ dV.
\end{align}
Now applying the Gaussian rule of integration on the second integral over the
surface, the resulting force in per unit volume is
\begin{align}
    \int_V \left(\rho \mathbf{F} - \nabla P\right)\ dV.
\end{align}
The acceleration of the fluid particles is given by $\frac{D\mathbf{u}}{Dt}$,
and thus the total force per unit volume on the other hand is
\begin{align}
    \int_V \rho \frac{D\mathbf{u}}{Dt}\ dV =
    \int_V \left(\rho \mathbf{F} - \nabla P\right)\ dV.
\end{align}
Newton's Second Law for a fluid in an Volume is essentially saying that the
rate of change of momentum of the fluid in the fixed volume $V$, which is the particle
acceleration is the resulting force acting on V together with the rate of
flow of momentum across the surface $S$ into the volume $V$. Hence we arrive
at the \textbf{Euler's Equation(s) of Motion}
\begin{align}
    \frac{D\mathbf{u}}{Dt} = \left(\frac{\partial \mathbf{u}}{\partial t}
    (\mathbf{u}\nabla)\mathbf{u}\right)  =
    -\frac{1}{\rho}\nabla P + \mathbf{F}.
\end{align}
As a side note we have mentioned that there is another contribution if the
fluid is viscid. Indeed there is a tangential force due to the velocity
gradient, which into introduces the additional term
\begin{align}
    \mu \nabla^2 \mathbf{u}, \qquad
    \mu = \text{viscosity of the Fluid}.
\end{align}
Thereby the equations become
\begin{align}
    \rho\frac{D\mathbf{u}}{Dt}
    =  -\nabla P + \rho \mathbf{F} + \mu \nabla^2 \mathbf{u}.
\end{align}

For now we have separated two simplifications, that define an
\textbf{idealized/perfect fluid}
\begin{enumerate}
    \item \textbf{incompressible} $\qquad \mu=0$
    \item  \textbf{inviscid} $\quad \rho = \text{const.},\ \nabla \mathbf{u}=
        0$
\end{enumerate}
\subsection{Vorticity and irrotational Flow}
The curl of the velocity field $\mathbf{\omega} = \nabla \times \mathbf{u}$
of a fluid (i.e. the vorticity), describes a spinning motion of the fluid
near a position $\mathbf{x}$ at time $t$. The vorticity is an important
property of a fluid, flows or regions of flows where $\mathbf{\omega}=0$ are
\textbf{irrotational}, and thus can be modeled and analyzed following well
known routine methods. Even though real flows are rarely irrotational
anywhere (!), in water wave theory wave problems, from the classical aspect
of vorticity have a minor contribution. Hence we can assume irrotational flow
modeling water waves. To arrive at the vorticity in the equations of motions
derived in the last section we resort to a differential identity derived in appendix
\ref{appendix:diff identity}, which gives for the material derivative
\begin{align}
    \frac{D\mathbf{u}}{Dt} = \frac{\partial \mathbf{u}}{\partial t}
    \nabla(\frac{1}{2}\mathbf{u}\mathbf{u)}
    - \left( \mathbf{u}\times (\nabla \times  \mathbf{u} \right).
\end{align}
Thus the equations of motion become
\begin{align}
    \frac{\partial \mathbf{u}}{\partial t} + \nabla\left(
    \frac{1}{2}\mathbf{u}\mathbf{u} + \frac{P}{\rho} + \Omega \right)
    = \mathbf{u} \times  \mathbf{\omega},
\end{align}
where $\Omega$ is the force potential per
unite mass given by $\mathbf{F} = -\nabla \Omega$.

At this point we may differentiate between \textbf{stead and unsteady flow}.
For \textbf{Steady Flow} we assume that $\mathbf{u}, P$ and $\Omega$ are time
independent, thus we get
\begin{align}
      \nabla\left( \frac{1}{2}\mathbf{u}\mathbf{u} + \frac{P}{\rho} + \Omega
      \right)  = \mathbf{u} \times  \mathbf{\omega}.
\end{align}
It is general knowledge that the gradient of a function $\nabla f$ is
perpendicular the level sets of $f(\mathbf{x})$, where $f(\mathbf{x}) =
\text{const.}$. Thus $\mathbf{u} \times  \mathbf{\omega}$ is orthogonal to
the surfaces  where
\begin{align} \label{eq:bernoulli}
    \frac{1}{2}\mathbf{u}\mathbf{u} + \frac{P}{\rho} + \Omega =
    \text{const.},
\end{align}
The above equation is called \textbf{Bernoulli's Equation}.

Secondly \textbf{Unsteady Flow} but irrotational (+ incompressible), first of
all gives us the condition for the existence of a velocity potential $\phi$
in the sense
\begin{align}
    \mathbf{\omega} = \nabla \times  \mathbf{u} = 0  \quad \Rightarrow \quad
    \mathbf{u} = \nabla \phi,
\end{align}
where $\phi$ needs to satisfy the Laplace equation
\begin{align}
    \Delta \phi = 0.
\end{align}
According to the Theorem of Schwartz we may exchange $\frac{\partial
}{\partial t}$ and $\nabla$, giving us an expression for the material
derivative
\begin{align}
    \nabla\left( \frac{\partial \phi}{\partial t} +\frac{1}{2}
    \mathbf{u}\mathbf{u} + \frac{P}{\rho}  + \Omega \right) = 0
\end{align}
Thus the expression differentiated by the $\nabla$ operator is an arbitrary
function $f(\mathbf{x}, t)$, writing
\begin{align}
     \frac{\partial \phi}{\partial t} +\frac{1}{2}
    \mathbf{u}\mathbf{u} + \frac{P}{\rho}  + \Omega = f(\mathbf{x}, t).
\end{align}
The function $f(\mathbf{x}, t)$ can be removed by gauge transformation of
$\phi \rightarrow \phi + \int f(\mathbf{x}, t)\ dt$, never the less this is
not further discussed and left to the reader in the reference.
\subsection{Boundary Conditions for water waves}
The boundary conditions for water-wave problems vary, generally on the
simplification we undertake. At the surface, called the free surface as in
free from the velocity conditions, we have the atmospheric stress on the
fluid. The stress component would again have a viscid component, this however
is only relevant when modeling surface wind, in this review we model the
fluid as unaffectedly and within reason as inviscid. The atmosphere employs
only a pressure on the surface, this pressure is taken to be the atmospheric
pressure, dependent on time and point in space. Thereby  any surface tension
effects can also include a scenario at a curved surface (e.g. wave), giving
rise to the pressure difference across the surface. A more precise
description would use Thermodynamics to derive boundary conditions coupling
water surface and the air above it, yet the density component of air
compared to that of water makes our ansatz viable. The described conditions
are called the \textbf{dynamic conditions}

An additional condition revolves around the fluid particles on the moving
surface, called the \textbf{kinematic condition}. This condition bounds
the vertical velocity component on the surface.

The logical step now is to define boundary conditions on the bod of the
fluid, i.e. the bottom. If the viscid case bottom is impermeable, we a no
slip condition to all fluid particles $\mathbf{u}_\text{bottom}= 0$. If we
assume that the fluid is inviscid then the bottom becomes a surface of the
fluid in the sense that the fluid particles in contact with the bed move in
the surface, we more or less mirror the kinematic condition of the surface.
For many problems the condition is going to vary, in most cases the bottom
will be rigid and fixed not necessarily horizontal. This condition is simply
called the \textbf{bottom condition}.
\subsubsection{Kinematic Condition}
Obtaining the free surface is the primary objective in the theory of modeling
water waves, represented by
\begin{align}
    z = h(\mathbf{x}_\perp, t),
\end{align}
where $\mathbf{x}_\perp = (x, y)$ in Cartesian, or $\mathbf{x}_\perp = (r,
\theta)$ in cylindrical coordinates. A surfaces that moves with the fluid,
always contains the same fluid particles, described as
\begin{align}
    \frac{D}{Dt}\left(z - h(\mathbf{x}_\perp, t  \right) = 0.
\end{align}
Upon expanding the derivative we get
\begin{align}
    \frac{Dz}{Dt} - \frac{Dh}{Dt}
    &= \frac{\partial z}{\partial t}+
    (\mathbf{u}\nabla)z - \frac{\partial h}{\partial t} -(\mathbf{u}\nabla)\\
    &= w - \left(h_t - (\mathbf{u}_\perp \nabla_\perp) h\right) = 0,
\end{align}
where the subscript $\perp$ describes the components with regard to
$\mathbf{x}_\perp$. The \textbf{kinematic condition} reads
\begin{align}
    w = h_t - (\mathbf{u}_\perp \nabla_\perp) h \qquad \text{on}\;\;
    z=h(\mathbf{u}_\perp, t).
\end{align}

\subsubsection{Dynamic Condition}
As described in the prescript of this section, the case of an inviscid fluid,
requires that only the pressure $P$ needs to be described on the free surface
$z = h(\mathbf{x}_\perp, t)$. Assuming incompressible, irrotational,
unsteady flow and setting $P=P_a$ for atmospheric pressure and $\Omega =
g\cdot z$ for the force per unit mass potential the equations of motion are
\begin{align}
    \frac{\partial \phi}{\partial t} +\frac{1}{2}\mathbf{u}\mathbf{u}
    + P_\frac{a}{\rho}+gh = f(t) \qquad \text{on}\;\; on z=h.
\end{align}
Somewhere $\|\mathbf{x}_\perp\| \rightarrow \infty$ the fluid reaches
equilibrium and is thereby stationary, thereby has no motion and the pressure
is $P=P_a$ and the surface is a constant $h = h_0$ $f(t)$ is
\begin{align}
    f(t) = \frac{P_a}{\rho}+gh_0.
\end{align}
The simplest description for the \textbf{dynamic condition} may be written as
\begin{align}
    \frac{\partial \phi}{\partial t}
    +\frac{1}{2}\mathbf{u}\mathbf{u}+g(h-h_0) = 0 \qquad \text{on}\;\; z=h.
\end{align}

Regarding the pressure difference on a curved surface, we may expand the
dynamic condition by introducing the pressure difference known as the
\textbf{Young-Laplace Equation}
\begin{align}
    \Delta P = \frac{\Gamma}{R},
\end{align}
where $\Gamma>0$ is the coefficient of surface tension and $\frac{1}{R}$ is
the curvature representing an implicit function, in our case the implicit
function is $z - h(\mathbf{x}_\perp, t)$ for fixed time. The curvature in
Cartesian coordinates takes the form
\begin{align}
    \frac{1}{R} = \frac{(1+h_y^2)h_{x x}+(1+h_y^2)h_{yy} -
    2h_xh_yh_{xy}}{\left( h_x^2+h_y^2+1 \right)^{\frac{3}{2}} },
\end{align}
the derivation is precisely described in \ref{appendix:curvature}



\subsubsection{The Bottom Condition}
The representation for the bottom is
\begin{align}
   z = b(\mathbf{x}_\perp, t),
\end{align}
where the fluid surface needs to satisfy
\begin{align}
    \frac{D}{Dt} \left(z - b(\mathbf{x}_\perp) \right)  = 0.
\end{align}
Hence we arrive at the bottom boundary conditions
\begin{align}
    w = b_t + (\mathbf{u}_\perp \nabla_\perp)b \qquad \text{on}\;\; z=b ,
\end{align}
where $b(\mathbf{x}_\perp, t)$ is already known for most water wave
problems. If we consider a stationary bottom then the time derivative
vanishes, leaving us with the following condition
\begin{align}
    w = (\mathbf{u}_\perp \nabla_\perp)b \qquad \text{on}\;\; z=b
\end{align}


\subsubsection{Integrated Mass Condition}
In this section we want to combine the kinematics of both the free and the
bottom surface with the mass conservation equation on the perpendicular
components
\begin{align}
    \nabla \mathbf{u} = \nabla_\perp \mathbf{u}_\perp + w_z = 0 .
\end{align}
Integrating the above expression from bottom to surface, i.e. from
$z=b(\mathbf{x}_\perp,t)$ to $z = h (\mathbf{x},t)$ gives
\begin{align}
    \int_b^h \nabla_\perp \mathbf{u}_\perp\ dz + w\bigg|_{z=b}^{z=h} = 0,
\end{align}
where we insert the conditions on the free surface and on the bottom surface
\begin{align}
    w &= h_t + (\mathbf{u}_{\perp \text{s}} \nabla_\perp) h \quad
    \text{on}\;\; z = h\\
    w &= b_t + (\mathbf{u}_{\perp \text{b}} \nabla_\perp) h \quad
    \text{on}\;\; z =b,
\end{align}
with the subscript $s$ and $b$ indicating the evaluation of a quantity
on the free surface and the bottom surface respectively. Inserting the
boundary conditions we get
\begin{align}
    \int_b^h \nabla_\perp \mathbf{u}_\perp
    + h_t + (\mathbf{u}_{\perp \text{s}} \nabla_\perp) h
    - b_t - (\mathbf{u}_{\perp \text{b}} \nabla_\perp) b= 0.
\end{align}
To simplify the equation we resort again to the Leibniz Rule of Integration
\begin{align}
     \int_b^h \nabla_\perp\mathbf{u}_\perp =
    \nabla_\perp \int_b^h \mathbf{u}_\perp\ dz - (\mathbf{u}_{\perp \text{s}}
    \nabla_\perp)h - (\mathbf{u}_{\perp \text{b}})b.
\end{align}
As a consequence the \textbf{Integrated Mass Condition} is given by
\begin{align}
    \nabla_\perp \int_b^h \mathbf{u}_\perp\ dz  + \underbrace{h_t -
    b_t}_{=d_t} = 0.
\end{align}
\subsection{Energy Equation}
To derive the energy equation we start off with Euler's Equation of Motion
\begin{align}
  \mathbf{u} _t + \nabla
  (\frac{1}{2}\mathbf{u}\mathbf{u}+\frac{P}{\rho}+\Omega) = \mathbf{u}\times
  \mathbf{w},
\end{align}
multiplying the equation with $\mathbf{u}$ we get
\begin{align}
  &\mathbf{u}\mathbf{u} _t \label{eq:energy1} \\
  &+(\mathbf{u}\nabla)(\frac{1}{2}\mathbf{u}\mathbf{u}+\frac{P}{\rho}+\Omega)\label{eq:energy2}\\
  &= \mathbf{u}(\mathbf{u}\times
  \mathbf{w})\label{eq:energy3}.
\end{align}
The first equation given in \ref{eq:energy1} can we rewritten using inverse
product rule of differentiation
\begin{align}
    \mathbf{u}\frac{\partial \mathbf{u}}{\partial t}
    &= \frac{\partial
    }{\partial t} (\mathbf{u}\mathbf{u}) - \frac{\partial \mathbf{u}}{\partial t}
    \mathbf{u} \\
    &= \frac{\partial
    }{\partial t} (\mathbf{u}\mathbf{u}) - \mathbf{u}\frac{\partial
    \mathbf{u}}{\partial t}\\
    \Rightarrow\quad & \mathbf{u} \frac{\partial \mathbf{u}}{\partial t}  =
    \frac{1}{2}\frac{\partial }{\partial t} (\mathbf{u}\mathbf{u}).
\end{align}
Then we may add
\begin{align}
    \left(\frac{1}{2} \mathbf{u}\mathbf{u}+\frac{P}{\rho} +\Omega  \right)
    \underbrace{(\nabla u)}_{=0} = 0,
\end{align}
to above not changing anything. Thereby getting
\begin{align}
    \frac{\partial }{\partial t} (\frac{1}{2}\mathbf{u}\mathbf{u})
    +(\mathbf{u}\nabla \mathbf{u})\left(
    \frac{1}{2}\mathbf{u}\mathbf{u}+\frac{P}{\rho} \right)
    +\left( \frac{1}{2}\mathbf{u}\mathbf{u}+\frac{P}{\rho} + \Omega \right)
    (\nabla \mathbf{u}) = 0.
\end{align}
Applying the product rule we can simplify
\begin{align}
    \frac{\partial }{\partial t} \left(\frac{1}{2}\mathbf{u}\mathbf{u}\right)
    +\nabla \left(\mathbf{u}\left(\mathbf{u}(
    \frac{1}{2}\mathbf{u}\mathbf{u}+\frac{P}{\rho}\right) \right) = 0,
\end{align}
additionally adding $\frac{\partial \Omega}{\partial t}  =0$ leads us to
\begin{align}
    \underbrace{\frac{\partial }{\partial t} \left(\frac{1}{2}\mathbf{u}\mathbf{u}
    +\Omega\right)}_{\text{change of total energy density}}
    +\underbrace{\nabla \left(\mathbf{u}\left(\mathbf{u}(
    \frac{1}{2}\mathbf{u}\mathbf{u}+\frac{P}{\rho}\right)
\right)}_{\text{energy flow of the velocity field}} = 0.\label{eq:energy}
\end{align}
This is called the \textbf{energy equation} and is a general result for a
inviscid and incompressible fluids, which we can apply to study water waves.
We start off with replacing $\nabla = \nabla_\perp + \frac{\partial }{\partial
z} $ and $\Omega = g z$ and multiplying by $\rho$, then our energy equation
in \ref{eq:energy} becomes
\begin{align}
    \frac{\partial }{\partial t}\left( \frac{1}{2}\rho\mathbf{u}\mathbf{u} + \rho
    g z\right)  + \nabla_\perp\left( \mathbf{u}_\perp\left(
    \frac{1}{2}\rho\mathbf{u}\mathbf{u}+P+\rho gz \right)  \right)
    \frac{\partial}{\partial z} \left( w\left(
    \frac{1}{2}\rho\mathbf{u}\mathbf{u}+P+\rho g z \right)  \right)  = 0.
\end{align}
Integrating from bottom to top, i.e. from bed to free surface gets us to
\begin{align}
    &\int_b^h\frac{\partial }{\partial t}\left( \frac{1}{2}\rho\mathbf{u}\mathbf{u} + \rho
    g z\right)\ dz  \label{eq:e-int1}\\
    &+ \int_b^h \nabla_\perp\left( \mathbf{u}_\perp\left(
    \frac{1}{2}\rho\mathbf{u}\mathbf{u}+P+\rho gz \right)  \right)\
    dz\label{eq:e-int2}\\
    &+ \left(\frac{\partial}{\partial z} \left( w\left(
    \frac{1}{2}\rho\mathbf{u}\mathbf{u}+P+\rho g z \right)
\right)\right)\Bigg|_b^h \label{eq:e-int3}
    = 0.
\end{align}
For equation \ref{eq:e-int1} we use Leibniz Rule of Integration, leaving us
with
\begin{align}
    \int_b^h\frac{\partial }{\partial t}\left( \frac{1}{2}\rho\mathbf{u}\mathbf{u} + \rho
    g z\right)\ dz
    &= \frac{\partial }{\partial t} \int_b^h
    \frac{1}{2}\rho\mathbf{u}\mathbf{u} + \rho gz \ dz\\
    &+ \left( \frac{1}{2}\rho \mathbf{u}_s \mathbf{u}_s  + \rho g h \right)
    h_t\\
    &- \left( \frac{1}{2}\rho \mathbf{u}_b \mathbf{u}_b  + \rho g b \right)
    b_t
\end{align}
For equation \ref{eq:e-int2} we again take note of the Leibniz Rule of
Integration, getting
\begin{align}
    \int_b^h \nabla_\perp\left( \mathbf{u}_\perp\left(
    \frac{1}{2}\rho\mathbf{u}\mathbf{u}+P+\rho gz \right)  \right)\
    dz
    &= \nabla_\perp \int_b^h \mathbf{u}_\perp\left(
    \frac{1}{2}\rho\mathbf{u}\mathbf{u} + P + \rho g z \right) \ dz\\
    &- \left( \frac{1}{2}\rho \mathbf{u}_s\mathbf{u}_s + P + \rho g h \right)
    \left( \mathbf{u}_{\perp s} \nabla_\perp \right) h\\
    &+\left( \frac{1}{2}\rho \mathbf{u}_b\mathbf{u}_b + P + \rho g b \right)
    \left( \mathbf{u}_{\perp b} \nabla_\perp \right) b
\end{align}
Thereby transforming our equation into
\begin{align}
    \frac{\partial }{\partial t} \underbrace{\int_b^h \frac{1}{2}\rho
    \mathbf{u}\mathbf{u}+\rho g z\ dz}_{=:\mathcal{E}}
    + \nabla_\perp&\underbrace{\int_b^h
    \mathbf{u}_\perp\left( \frac{1}{2}\rho\mathbf{u}\mathbf{u} + \rho g z
\right)\ dz}_{:=\mathcal{F}}
+ \underbrace{P_s h_t - P_b b_t}_{:=\mathcal{P}} = 0\\
\nonumber\\
    &\frac{\partial \mathcal{E}}{\partial t}
    + \nabla_\perp \mathcal{F} + \mathcal{P} = 0,
\end{align}
where $\mathcal{E}$ represents the energy in the flow per unit horizontal
area, since we are integrating from bed to free surface. Where $\mathcal{F}$
is the horizontal energy flux vector and lastly $\mathcal{P} = P_s h_t -
P_b b_t$ is the net energy input due to the pressure forces doing work on the
upper and lower boundaries, i.e. bottom and free surface of the fluid.
Assuming stationary rigid bottom condition and constant surface pressure, we
can set $P_s=0$, such that $\mathcal{P} =0$ leaving us with the equation
\begin{align}
    \frac{\partial \mathcal{E}}{\partial t}
    + \nabla_\perp \mathcal{F} = 0.
\end{align}
We note that the assumption $P_s=0$ is only possible if the coefficient of
surface tension is set to 0, which usually is not the case.
